---
title: |
  | Additional Material for Creating Matrices 
  | using Data Frames
author: "Sudhakar Kumar"
date: "1 March 2019"
output: pdf_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Creating vectors in R 
A **vector** is a sequence of data elements of the same type. In **R** language, vectors are created using the `c` function. For example, suppose we want to create a vector with its elements as `(1,2,3,4)`. Then, we use the following command 
```{r echo=TRUE}
c(1,2,3,4)
```
We can also use the command as follows 
```{r echo=TRUE}
c(1:4)
```
If we want to create a vector of consecutive numbers, the `:` operator comes very handy. Hence, for creating a vector with elements from 1 to 4, we can type the command as follows
```{r echo=TRUE}
vec <- 1:4
vec
```
Since, a vector must have data elements of the same type, this function `c` will try and coerce elements to the same type, if they are different.

Coercion is from lower to higher types from logical to integer to double to character [@DataMentor]. For example, if we create a vector with data elements of different type as given below
```{r echo=TRUE}
vec <- c(1, 5.4, TRUE, "hello")
vec
```
As we see that we have inserted different data types like numeric, logical and character in this vector `vec`. However, **R** language will coerce all the elements to character type. Let us find out the data type of our vector `vec` (using the `class` or `typeof` function). 
```{r echo=TRUE}
class(vec)
```
More complex sequences can be created using the `seq` function, like defining the number of points in an interval, or the step size. For example, we want to create a vector with elements from 0 to 5 with a step size of 0.5. For this, we type the command as given below
```{r echo=TRUE}
seq(from = 0, to = 5, by=0.5)
```
# Creating matrices using vectors 
Matrix can be created using the `matrix` function. Dimension of the matrix can be defined by passing appropriate value for arguments `nrow` and `ncol`. Providing values for both dimension is not necessary. If one of the dimension is provided, the other is inferred from length of the data. 
Suppose we want to create a $3\times3$ matrix with numbers from 1 to 9, we type the command as follows 
```{r echo=TRUE}
matrix(data = 1:9, nrow = 3, ncol = 3)
```
Here, we can see that the numbers from 1 to 9 has been arranged column wise. If we want the numbers (1 to 9) to be arranged in row wise format, we type the command as given below
```{r echo=TRUE}
matrix(data = 1:9, nrow = 3, ncol = 3, byrow = TRUE)
```

Consider two different vectors as given below
```{r echo=TRUE}
vec1 <- c(9:12)
vec2 <- c(13:16)
vec1; vec2
```

Now, we want to create a $4\times2$ matrix using these two vector `vec1` and `vec2`. For this, we will first create a single vector by joining these two vectors, as given below
```{r echo=TRUE}
vec_final <- c(vec1, vec2)
vec_final
```
Then, we use the `matrix` function and apply the arguments as given below:

* **data** = `vec_final` 
* **nrow** = 4
* **ncol** = 2

So, we type the command as given below
```{r}
mat <- matrix(data = vec_final, nrow = 4, ncol = 2)
mat
```


# References

