---
title: ' Additional Material on Data Types and Factors ^[For any suggestions/comments, please write to contact-r@fossee.in]'
output:
  pdf_document: default
  html_document:
    df_print: paged
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data types in R 
In R programming language, the variables are not declared as some data type unlike other programming languages e.g. **C** and **Java**. The variables are assigned with R-objects and the data type of the R-object becomes the data type of the variable [@tutpoint]. Hence, everything in R is an object. There are many types of R-objects. The frequently used ones are as given below: 

* **Vector**
* **List**
* **Matrix**
* **Array**
* **Factor**
* **Data Frame**

The simplest of these objects is the **vector** object. R language has five frequently used atomic vector types [@carpent]: 

* **Logical** -- `TRUE`, `FALSE`
* **Integer** -- `2L`, `4L`, etc. (the `L` tells **R** to store this as an integer)
* **Numeric** -- `5`, `12.5`, etc.
* **Complex** -- `1 + 4i`, `2 + i`, etc.
* **Character** -- `"Hello"`, `"R"`, etc. 

By *atomic*, we mean the vector only holds data of a single type. We will learn how to declare these atomic vector types. 

A logical type includes two elements namely, `TRUE` and `FALSE`.

```{r echo=TRUE}
testData <- FALSE
print(class(testData))
```

In order to declare an integer in R language, we need to add an `L` suffix. Alternatively, one can also use `as.integer` function. 

```{r echo=TRUE}
testData <- 5L
print(class(testData))
```

```{r echo=TRUE}
testData <- as.integer(5)
print(class(testData))
```

The numeric vector type is very simple to declare. We can assign the desired number directly to an object.

```{r echo=TRUE}
testData <- 12.5
print(class(testData))
```

```{r echo=TRUE}
testData <- 12
print(class(testData))
```

A complex number has two parts: real and imaginary. The imaginary part is written with an `i` suffix. 

```{r echo=TRUE}
testData <- 5 + 4i
print(class(testData))
```

A character is written within double quotes and then assigned to an object. 

```{r echo=TRUE}
testData <- "R Programming Language"
print(class(testData))
```

There is one more vector type called `raw`, which is intended to hold raw bytes. For declaring an object of raw type, we use `charToRaw` function. 

```{r echo=TRUE}
testData <- charToRaw("Hello")
print(testData)
print(class(testData))
```

`R` provides many built-in functions to examine features of vectors and other objects. For example, we can find the length of a vector by using `length` function. 

```{r echo=TRUE}
testData <- seq(from = 1, to = 10, by = 0.5)
paste("The length of testData is", length(testData))
```

Also, `class` and `typeof` functions are used to find the data type of the object. 

# Creating vectors in R 
One can create an empty vector using `vector` function. For example, we want to create an empty character vector of length 3. 

```{r echo=TRUE}
vector("character", length = 3)
```

Similarly, we can create a numeric vector of length 5 by modifying the values of arguments present in `vector` function. 

```{r echo=TRUE}
vector("numeric", length = 5)
```



# References 
