---
title: "Additional Material for Lists and its Operations"
author: "Sudhakar Kumar"
date: "25 April 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lists in R 
A list is an R object. It has components of mixed data types like 

* strings
* numbers
* vectors or 
* some other list inside it 

A list can also contain a matrix or a function as its elements. 

# Creating a list in R 
For example, we will create a list containing a vector, a matrix and one built-in dataset. First, we need to declare these three components. As `iris` dataset is having 150 rows, we will consider its seven rows only.  

```{r echo=TRUE}
myVector <- c(1:5)
myMatrix <- matrix(1:15, nrow = 5, ncol = 3, byrow = TRUE)
myDataSet <- iris[c(1:2,51:52,100:102),]
```

Now, we use `list` function to create the list and `names` function to add names to the elements of our list. 

```{r echo=TRUE}
myList <- list(myVector, myMatrix, myDataSet)
names(myList) <- c("vector", "matrix", "dataset")
print(myList)
```

# Accessing the elements of a list 

In case of named list, we can retrieve the elements by their names. 

```{r echo=TRUE}
myList$dataset
```

Also, the elements of a list can be retrieved by using the single square bracket `[]` operator.

```{r echo=TRUE}
myList[3]
```

In order to reference a list member directly, we have to use the double square bracket `[[]]` operator. 

```{r echo=TRUE}
myList[[3]]
```

Thus, we can observe that the difference between single square bracket and double square brackets is that with double square brackets, the name of the element (`dataset`) is not displayed. According to a discussion in the forum of   **Analytics Vidhya**, we need `[[` (double square brackets) when working with lists. This is because when `[` (single square bracket) is applied to a list, it always returns a list: it never gives us the contents of the list. To get the contents, we need `[[` (double square brackets).  This can also be verified by using `class` function. 

```{r}
class(myList$dataset); class(myList[3]); class(myList[[3]])
```

