---
title: "| Additional Material for Plotting Histograms \n| and Pie Chart\n"
author: "Sudhakar Kumar"
date: "5 May 2019"
output:
  pdf_document: default
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data frame for analysis 
A data frame [@Mine] comprising information on movies has been used. It contains 651 observations (rows), each representing a movie, and 32 variables (columns). This data frame can be loaded in **R** either by using `read.csv` function or by `load` function. 

```{r echo=TRUE}
load(url("https://stat.duke.edu/~mc301/data/movies.Rdata"))
dim(movies)
```

# Handling the missing values in data frame
There is an object named `runtime` in `movies` data frame. `runtime` denotes the runtime of movie (in minutes). We can find the range of `runtime` in `movies` by using `range` function. `range` function returns a vector containing the minimum and maximum of the given argument(s).

```{r echo=TRUE}
range(movies$runtime)
```

**R** returns `NA` and `NA` as minimum and maximum runtime of `movies`. It means that `runtime` object contains `NA`. 

`NA` stands for **not available**. `NA` [@Edwin]  is a placeholder for a missing value. All basic operations in **R** handle `NA` without crashing and mostly return `NA` as an answer whenever one of the input arguments is `NA`. To know, the range of `runtime` in movies, we need to remove the **NAs**. For this, we set the argument `na.rm` to be `TRUE`. 

```{r echo=TRUE}
minMax <- range(movies$runtime, na.rm = TRUE)
paste("Minimum runtime is", minMax[1], "minutes.")
paste("Maximum runtime is", minMax[2], "minutes.")
```

# References 
