---
title: "| Additional Material for Plotting Bar Charts and \n| Scatter Plot\n"
#author: "Sudhakar Kumar"
#date: "29 May 2019"
output: pdf_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data frame for analysis
We will use a data frame [@Mine] comprising information on **movies**. It contains 651 observations (rows), each representing a movie, and 32 variables (columns). Let us load this data frame by using `load` function.

```{r echo=TRUE}
load(url("https://stat.duke.edu/~mc301/data/movies.Rdata"))
paste("Information on", dim(movies)[1], "movies loaded.")
```

In **movies** data frame, there is one variable named `imdb_rating`. It contains the IMDB rating (on a scale of 0 to 10) for each of the 651 movies. For example, the movies titled **The Godfather, Part II** and **Disaster Movie** have lowest and highest IMDB rating, respectively. This can also be verified by using `range` function.   

```{r echo=TRUE}
minMax <- range(movies$imdb_rating)
paste(movies$title[movies$imdb_rating==minMax[1]], "has lowest IMDB rating of", minMax[1])
paste(movies$title[movies$imdb_rating==minMax[2]], "has highest IMDB rating of", minMax[2])
```

# Plotting a bar chart
A **bar chart** represents data in rectangular bars with length of the bar proportional to the value of the variable [@tutbar]. **R** uses the function `barplot` to create **bar charts**. We can draw both vertical and horizontal bars in the **bar chart** using **R**. The basic syntax to create a **bar chart** in **R** is: 

`barplot(data, xlab, ylab, main, names.arg, col)`

Following is the description of the parameters used:

* `data` is either a vector or matrix of values describing the bars which make up the plot.
* `xlab` and `ylab` are the labels for x and y axes, respectively.
* `main` is the title of the bar chart.
* `names.arg` is a  vector of names to be plotted below each bar or group of bars.
* `col` is used to give colors to the bars in the graph.

We will create a **bar chart** from the variable `imdb_rating`. For the sake of simplicity, we consider only the first 20 observations of **movies** data frame. 

```{r echo=TRUE}
moviesSub <- movies[1:20,]
barplot(moviesSub$imdb_rating,
        ylab = "IMDB rating",
        col = "blue",
        ylim = c(0,10),
        main = "Movies' IMDB rating",
        names.arg = moviesSub$title,
        las = 2)
```

In this **bar chart**, we can see that the titles of movies are present on the X-axis. However, the longer titles are being truncated. This can be fixed either by defining the plotting area or by adding more functions followed by `barplot` function. 

#Adjusting labels of a bar chart by using par function 
First, we will apply the former method by using `par` function. According to the documentation of **R** programming language, `par` can be used to set or query graphical parameters. Parameters can be set by specifying them as arguments to `par` in `tag = value` form, or by passing them as a list of tagged values. Here, we pass a `tag` named `mar`, which refers to margin. It is a numerical vector of the form `c(bottom, left, top, right)` which gives the number of lines of margin to be specified on the four sides of the plot.

```{r echo=TRUE}
par(mar=c(11,5,5,2)) # c(bottom, left, top, right)
barplot(moviesSub$imdb_rating,
        ylab = "IMDB rating",
        col = "blue",
        ylim = c(0,10),
        main = "Movies' IMDB rating",
        names.arg = moviesSub$title,
        las = 2)
```

Now, the names are not being truncated, as it is evident from the graph plotted above.

#Adjusting labels of a bar chart by using text function 
Here, we will add one more function named `text` after the `barplot` function. According to the documentation of **R** programming language, `text` draws the strings given in the vector labels at the coordinates given by `x` and `y`. The required coordinates can be found out by trial and error method. We will pass the following arguments in this function:

* **coordinates** (x and y) -- numeric vectors of coordinates where the text labels should be written. 
* **labels** -- a character vector or expression specifying the text to be written.
* **xpd** -- used to enable things to be drawn outside the plot region.
* **srt** -- used to rotate the text by an angle.

```{r echo=TRUE}
x <- barplot(moviesSub$imdb_rating, 
             ylab = "IMDB rating",
             ylim = c(0,10),
             main = "Movies' IMDB rating",
             col = "blue")
labs <- moviesSub$title
text(x = x - 0.95, 
     y = -1.25, 
     labs, 
     xpd = TRUE, 
     srt = 50)
```

# References 